---
title: "Linear Regression Report"
output:
  html_document:
    df_print: paged
    keep_md: yes
  pdf_document: default
     
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.width = 6, fig.asp = 0.618)
#libraries load
library(liver)
library(readr)
library(dplyr)
library(ggplot2)
library(moments)
library(readxl)
library(plot3D)
library(car)
library(corrplot)
library(lmtest)
```


## Data
### Independent Variables:
* Abs280
* Abs340
* Abs420
* Abs520
* Abs620
* Antocianins
* Catechins
* Polyphenol
* Color

### Dependent Variables:
* TT-IPT
* TT-MCP


```{r message=FALSE, warning=FALSE, include=FALSE}

data_taninos<- read_xlsx("/Users/leonardo/Documents/Vinos/dataset_taninos2023_2.xlsx", sheet = 'dataset')

```

## Normal Distribution and Outliers
\
```{r message=FALSE, warning=FALSE, include=FALSE}
#Analysis if distribution of each variable

#function to print the normality of the data. Import 'moments' library for the analysis.
swnormality_test <- function(ind_var) { 
  stest <- shapiro.test(ind_var)
  p_val <- stest[2]
  sprintf ("The p-value is %s:", p_val)
}
swnormality_test(data_taninos$Abs280)
swnormality_test(data_taninos$Abs340)
swnormality_test(data_taninos$Abs420)
swnormality_test(data_taninos$Abs520)
swnormality_test(data_taninos$Abs620)
swnormality_test(data_taninos$Antocianins)
swnormality_test(data_taninos$Catechins)
swnormality_test(data_taninos$Polyphenols)
swnormality_test(data_taninos$Color)
swnormality_test(data_taninos$TPI)
swnormality_test(data_taninos$TT_TPI)
swnormality_test(data_taninos$TT_MCP)

```


Graphical overview of normal data and outliers, complemented with a Shapiro-Wilkman normality test

```{r fig.show='hold', message=FALSE, warning=FALSE, include=FALSE, out.width="32%"}
# function that generates a histogram, boxplot and quantile plot to overview the distribution of the variables
norm_explore <- function(dat_frm, idp_vrbl) {
  
  histgrm <-   ggplot(dat_frm) +
    geom_histogram(aes_string(x = idp_vrbl), fill = "royalblue", bins = 10)+
    labs(title = idp_vrbl) +
    theme(aspect.ratio = 1)
  
  bx_plt <- ggplot(dat_frm, aes_string(idp_vrbl, idp_vrbl)) +
    geom_boxplot()+
    theme(aspect.ratio = 1)
  
  qq <- ggplot(dat_frm, aes_string(sample = idp_vrbl))+
    stat_qq() + stat_qq_line()+
    labs(title = idp_vrbl)+
    theme(aspect.ratio = 1)
 
  print (histgrm)
  print (bx_plt)
  print (qq)
}

norm_explore(data_taninos, "Abs280")
norm_explore(data_taninos, "Abs340")
norm_explore(data_taninos, "Abs420")
norm_explore(data_taninos, "Abs520")
norm_explore(data_taninos, "Abs620")
norm_explore(data_taninos, "Antocianins")
norm_explore(data_taninos, "Catechins")
norm_explore(data_taninos, "Polyphenols")
norm_explore(data_taninos, "Color")
norm_explore(data_taninos, "TPI")
norm_explore(data_taninos, "TT_TPI")
norm_explore(data_taninos, "TT_MCP")



```
\
According to the graphs and to the SW test, all the variables present a normal distribution. *Outliers based on the Polyphenol variable have been removed. 

The p-value for Abs340 is 0.0176, this data does not have a normal distribution. Thus, it is discarded from the analysis.\
\

## Correlation Assumption 
Correlation test between independent variables and TPI. 

```{r warning=FALSE, include=FALSE}
coefficient <- cor.test(data_taninos$Abs280, data_taninos$TT_TPI)
coefficient$estimate

coefficient <- cor.test(data_taninos$Abs340, data_taninos$TT_TPI)
coefficient$estimate

coefficient <- cor.test(data_taninos$Abs420, data_taninos$TT_TPI)
coefficient$estimate

coefficient <- cor.test(data_taninos$Abs520, data_taninos$TT_TPI)
coefficient$estimate

coefficient <- cor.test(data_taninos$Abs620, data_taninos$TT_TPI)
coefficient$estimate

coefficient <- cor.test(data_taninos$Antocianins, data_taninos$TT_TPI)
coefficient$estimate

coefficient <- cor.test(data_taninos$Catechins, data_taninos$TT_TPI)
coefficient$estimate

coefficient <- cor.test(data_taninos$Polyphenols, data_taninos$TT_TPI)
coefficient$estimate

coefficient <- cor.test(data_taninos$Color, data_taninos$TT_TPI)
coefficient$estimate


```
\
```{r Correlation Graph, warning=FALSE, dev = 'tiff', dpi= 900}
#Gráfica de correlacion entre todas las variables significativas

library(corrplot)

My_colors_p <- colorRampPalette(c("#74714E", "#A9C870" ,"#258EA6"))(20)

corrx <- cor(data_taninos[, c("TT_TPI","TT_MCP","Abs280", "Abs340", "Abs420", "Abs520", "Abs620","Catechins", "Antocianins", "Polyphenols", "Color")], method = "pearson")

colnames(corrx) <- c("TT-TPI","TT-MCP","Abs280", "Abs340", "Abs420", "Abs520", "Abs620","Catechins", "Anthocyanins", "Polyphenols", "Color")
rownames(corrx) <- c("TT-TPI","TT-MCP","Abs280", "Abs340", "Abs420", "Abs520", "Abs620","Catechins", "Anthocyanins", "Polyphenols", "Color")
plot_corr <- corrplot.mixed(corrx,
         color = "transparent",
         is.corr = FALSE,
         col.lim = c(-0.10,1.0),
         tl.pos = "lt",
         tl.col = "black",
         tl.cex = 0.7,
         number.cex = 0.6,
         lower.col = My_colors_p,
         upper.col = My_colors_p)

# Colors 
# #361134-purple; #726DA8 - lavender; ##861388- purple; pink #FFC6D9; purple #7E2E84
               
               
```

\
\
There is a high linear correlation for variables (r > 0.9)
+ Polyphenols -> 0.9670
\
There is a high linear correlation for variables (0.9 >r > 0.7):\
+ Abs280 -> 0.8574\
+ Color -> 0.8105\
+ Abs620 -> 0.7577\
+ Abs420 -> 0.7657\
+ Abs340 -> 0.7657\

There is moderate linear correlation for variables (0.7 > r > 0.5):\
+ Abs520 -> 0.5696\
+ Antocianins -> 0.5674\

There is a a low linear correlation for variable ( r < 0.3):\
+ Catechins -> 0.0636

---

## Linear Regression
\
```{r message=FALSE, warning=FALSE, include=FALSE}

data_znormalized <- zscore(data_taninos)

Data_Nrmlizd <- data_znormalized[, c("Polyphenols", "Abs280", "Color", "Abs420", "Abs620", "TPI", "TT_TPI", "TT_MCP")]

Normalized_data <- Data_Nrmlizd %>%
  bind_cols(data_taninos$`Sample code`)%>%
  rename(Sample_code = ...9)
```

Linear Regression with each variable:\
```{r message=FALSE, warning=FALSE, include=FALSE}
#function to generate linear regression graphs, residual vs fitted graph, q-q plot, outliers, and print out R-squared results for each model. 

Linear_Regression <- function(dataset, x, y){
  model_construction <- lm(dataset[[y]] ~ dataset[[x]], data = dataset)

  #linear regression model 
  model_graph <- ggplot(dataset, aes_string(x, y)) +
    geom_point()+
    geom_smooth(method = "lm") +
    geom_smooth(se = FALSE, color = "red")+
    theme_bw()
  print (model_graph)
  
  
  #Residuals vs fitted point graph
  resi_plot <- ggplot(model_construction, aes(x = .fitted, y = .resid)) +
    geom_point() +
    geom_hline(yintercept = 0) +
    labs(title='Residual (y)vs. Fitted Values(x) Plot', x=y, y=x)+
    geom_smooth(se = FALSE) +
    theme_bw()
  print (resi_plot)
  
  
  # Create Q-Q plot from residuals, to evaluate the assumption of ERRORS having a NORMAL DISTRIBUTION
  qq_plot_res <- ggplot(model_construction, aes(sample = .resid)) +
    stat_qq() + stat_qq_line()+
    theme(aspect.ratio = 1)+
    xlab("Theoretical Quantiles") +
    ylab("Sample Quantiles") +
    ggtitle("Normal Q-Q Plot of Residuals") +
    theme_bw()
  print(qq_plot_res)
  
  #Otra forma de identificar las observaciones que puedan ser outliers
  plot(x = model_construction$fitted.values, y = abs(rstudent(model_construction)),
main = "Absolute studentized residuals vs predicted values", pch = 20, col = "grey30")
abline(h = 3, col = "red")

  #'''model_residuals <- resid(model_construction)
  #plot(model_residuals,
  #ylab=y, xlab=x, main="Residuals Plot") 
  #abline(0, 0)'''
  
  model_predicts <- predict(model_construction)
  model_predicts <- residuals(model_construction)
  
  summary(model_construction)
  RSE_val <- signif(sigma(model_construction), digits = 4)
  Rsq_val <- signif(summary(model_construction)$r.squared, digits = 4)
  
RSS <- deviance(model_construction)
  
  print(paste('Model fit measurement for',x, ' is  R^2 = ',Rsq_val, 'and RSE=', RSE_val, 'with an RSS =',RSS))

}

 
Linear_Regression(data_taninos, "Abs280", "TT_TPI")
Linear_Regression(data_taninos, "Abs340", "TT_TPI")
Linear_Regression(data_taninos, "Abs420", "TT_TPI")
Linear_Regression(data_taninos, "Abs520", "TT_TPI")
Linear_Regression(data_taninos, "Abs620", "TT_TPI")
Linear_Regression(data_taninos, "Antocianins", "TT_TPI")
Linear_Regression(data_taninos, "Catechins", "TT_TPI")
Linear_Regression(data_taninos, "Polyphenols", "TT_TPI")
Linear_Regression(data_taninos, "Color", "TT_TPI")



```

The linear regression between **Polyphenols and TT-TPI** has a the best model fit scores, the values are:  
**R^2 =  0.933 and RSE= 71.11**.  \
This means, 93.5% of the variation is explained by TT-TPI. \
\
The Equation generated is: **TT-TPI = 1.07377xPolyphenols** \


## Multiple Linear Regression
\

#### Multiple linear regression using **TT-TPI** as dependat variable

```{r include=FALSE}
#Best predictors selection using AIC method
MLT_TT_TPI <- lm(TT_TPI ~ Abs420 +Abs620 +Antocianins +Catechins + Polyphenols + Color , data = data_taninos)
summary(MLT_TT_TPI)

#Best predictors selection using 
step(object = MLT_TT_TPI, direction = "both", trace = 1)

MLT_TT_TPI1 <- lm(TT_TPI ~ Abs620 + Polyphenols , data = data_taninos)
summary(MLT_TT_TPI1)

#Linear regression compared, resid vs fit

plot4 <- ggplot(data = data_taninos, aes(Abs620, MLT_TT_TPI1$residuals)) + geom_point() + geom_smooth(color = "firebrick") + geom_hline(yintercept = 0) + theme_bw()


plot7 <- ggplot(data = data_taninos, aes(Polyphenols, MLT_TT_TPI1$residuals)) + geom_point() + geom_smooth(color = "firebrick") + geom_hline(yintercept = 0) + theme_bw()

print(plot4)
print(plot7)

##QQ plot de residuals
qqnorm(MLT_TT_TPI1$residuals)
qqline(MLT_TT_TPI1$residuals)

## Normalidad SW
shapiro.test(MLT_TT_TPI1$residuals)

## Residuals distribution
MLR_resid <- ggplot(data = data_taninos, aes(MLT_TT_TPI1$fitted.values, MLT_TT_TPI1$residuals)) + geom_point() + geom_smooth(color = "firebrick", se = FALSE) + geom_hline(yintercept = 0) + theme_bw()
print(MLR_resid)

#Variance Inflation Factor (VIF): must be less than 5
vif(MLT_TT_TPI1)

#Autocorrelación
dwt(MLT_TT_TPI1, alternative = "two.sided")

##Atipic or influential points identification
plot(x = MLT_TT_TPI1$fitted.values, y = abs(rstudent(MLT_TT_TPI1)),
main = "Absolute studentized residuals vs predicted values", pch = 20, col = "grey30")
abline(h = 3, col = "red")

```


\
```{r include=FALSE}
#Manual Forward selection of best variables

my_MLR <- lm(TT_TPI ~ Polyphenols + Abs620 , data = data_taninos)
summary(my_MLR)
deviance(my_MLR)
vif(my_MLR)

my_MLR <- lm(TT_TPI ~ Polyphenols + Abs620 + Abs420, data = data_taninos)
summary(my_MLR)
deviance(my_MLR)

my_MLR <- lm(TT_TPI ~ Polyphenols + Abs620 + Antocianins , data = data_taninos)
summary(my_MLR)
deviance(my_MLR)

my_MLR <- lm(TT_TPI ~ Polyphenols + Abs620 + Catechins , data = data_taninos)
summary(my_MLR)
deviance(my_MLR)



```

\

The model that fits best to the data is given by the variables:  \
**TT-IPT ~ Polyphenols + Abs620** \
Residual standard error: 69.14\
Multiple R-squared:  0.9417\
The Equation generated is: **TT-IPT = 0.98801xPolyphenols + 108.13873xAbs620**


\
```{r warning=FALSE, include=FALSE}
#Graph for the MLR generated
z <- data_taninos$TT_TPI
x <- data_taninos$Polyphenols
y <- data_taninos$Abs620

fit <- lm(z ~ x + y)

# create a grid from the x and y values (min to max) and predict values for every point
# this will become the regression plane
grid.lines = 40
x.pred <- seq(min(x), max(x), length.out = grid.lines)
y.pred <- seq(min(y), max(y), length.out = grid.lines)
xy <- expand.grid( x = x.pred, y = y.pred)
z.pred <- matrix(predict(fit, newdata = xy), 
                 nrow = grid.lines, ncol = grid.lines)

# create the fitted points for droplines to the surface
fitpoints <- predict(fit)

# scatter plot with regression plane
scatter3D(x, y, z, pch = 19, cex = 1,colvar = NULL, col="#EA3323", 
          theta =18, phi = 30, bty= "b2",
          xlab = "Polyphenols", ylab = "Abs620", zlab = "TT-TPI",  
          surf = list(x = x.pred, y = y.pred, z = z.pred,  
                      facets = TRUE, fit = fitpoints, col=ramp.col (col = c("dodgerblue3","seagreen2"), n = 300, alpha=0.6), border="#74526C"), main = "TT-TPI")

```
\
```{r , include=FALSE}
Poly_color <- lm(TT_TPI ~ Polyphenols + Abs620, data = data_taninos) 
avPlots(Poly_color, main = "TT-TPI ~ Polyphenols + Abs620", pch =20, grid=F )

```
\